---
title: "Reproducible Document"
author: "Sofie H. G. Jakobsen,(this is yammel... a specefic syntact to format the qmd file)"
format: html
---

## Writing R code 

This is going to be text in the HTML file if u wanna make a code chunk you have to use:

"option+command+I"
                  
This will open the a code chunk:

```{r}
2+2
```

## The tasks from reading 15.6

To create an object we use: "<-"

```{r}
weight_kilos <- 100
weight_kilos
```


## The different data types and functions
```{r}
# Character vector
a1 <- c("a", "b", "c")
a1

# Logic vector
b1 <-c(TRUE, FALSE, FALSE)
b1

# Numeric vector
c1 <- c(1, 5, 6, 1.4)
c1

# Factor vector (special types of character vectors)
d1 <- factor(c("low", "high", "medium", "high"))
d1

#This is what a data frame looks like, if we look at the built-in dataset called airquality, which is a data frame object loaded by default when you start R
head(airquality)
```


## Understanding of libraries 

In the Packages tab you can find the packages you can activate using library. 
to get more packages u have to install them using install.packages(). 
Obs this should be done in the beginning of the document. 
we use packages from a system from the tidyverse. this includes a ocean of functions.
Tidyvers were installed as a part of the pre course material. 
Tidyvers is knwn as a meta packages. it loades 9 packages, very good for begining. 

Using "Setup" tells the code to run this chunch first

```{r setup}
library(tidyverse)
```


# Learning Markdown
# Header 1

Paragraph.

## Header 2

Paragraph.

### Header 3

Paragraph.

- item 1
- item 2
- item 3
1. item 1
2. item 2
3. item 3

**bold**

*italics*

super^script^

sub~script~



# Exercise 15.9: Practice using Markdown

## About me

-   Sofie H. G. Jakobsen
-   SDU

I'm a PhD Student in **Anne Loft's Group** at the *SDU*

## Simple code


```{r}
3*3
```

## setting up git
so for git you find the icon on the top bare it is a green + abnd red - over a G. from here you opent get by clicking commit. 
in there there is changes and history. in changes files with new changes will be shown and they can be commited with a comment describeing the change. a good trick is to add what file u add in the comment to make it very clear. 
In the history you can see the documets with the changes and the history of changes. 

some files u do not wish to changes like the html or any folder ending on _files. this can be added to git ignore so it wont be re uploaded everytime: 

```{r}
#usethis::use_git_ignore(c("*html", "*_files/"))

```



## Testing for Git


