---
title: "Reproducible Document"
author: "Sofie H. G. Jakobsen,(this is yammel... a specefic syntact to format the qmd file)"
format: html
---

## Writing R code 

This is going to be text in the HTML file if u wanna make a code chunk you have to use:

"option+command+I"
                  
This will open the a code chunk:

```{r}
2 + 2
```

## The tasks from reading 15.6

To create an object we use: "<-"

```{r}
weight_kilos <- 100
weight_kilos
```


## The different data types and functions
```{r}
# Character vector
a1 <- c("a", "b", "c")
a1

# Logic vector
b1 <- c(TRUE, FALSE, FALSE)
b1

# Numeric vector
c1 <- c(1, 5, 6, 1.4)
c1

# Factor vector (special types of character vectors)
d1 <- factor(c("low", "high", "medium", "high"))
d1

# This is what a data frame looks like, if we look at the built-in dataset called airquality, which is a data frame object loaded by default when you start R
head(airquality)
```


## Understanding of libraries 

In the Packages tab you can find the packages you can activate using library. 
to get more packages u have to install them using install.packages(). 
Obs this should be done in the beginning of the document. 
we use packages from a system from the tidyverse. this includes a ocean of functions.
Tidyvers were installed as a part of the pre course material. 
Tidyvers is knwn as a meta packages. it loades 9 packages, very good for begining. 

Using "Setup" tells the code to run this chunch first
option + I = | the line

```{r setup }
#| message: false
#| warning: false


library(tidyverse)
```


# Learning Markdown
# Header 1

Paragraph.

## Header 2

Paragraph.

### Header 3

Paragraph.

- item 1
- item 2
- item 3
1. item 1
2. item 2
3. item 3

**bold**

*italics*

super^script^

sub~script~



# Exercise 15.9: Practice using Markdown

## About me

-   Sofie H. G. Jakobsen
-   SDU

I'm a PhD Student in **Anne Loft's Group** at the *SDU*

## Simple code


```{r}
3 * 3
```

## setting up git
so for git you find the icon on the top bare it is a green + abnd red - over a G. from here you opent get by clicking commit. 
in there there is changes and history. in changes files with new changes will be shown and they can be commited with a comment describeing the change. a good trick is to add what file u add in the comment to make it very clear. 
In the history you can see the documets with the changes and the history of changes. 

some files u do not wish to changes like the html or any folder ending on _files. this can be added to git ignore so it wont be re uploaded everytime: 

```{r}
# usethis::use_git_ignore(c("*html", "*_files/"))
```



#### Testing for Git



# **Day 2** 
## **Importing data into R**

This data is from an openscience archive. we will be using this for this course. We look at insulin lvl. 
Do not run this every time! we only want to download it once
```{r}
# download.file(
#  url = "https://zenodo.org/records/4989220/files/Eliasson_data2.csv?download=1",
#  destfile = here::here("data/post-meal-insulin.csv")
# )


# Load the data into the enviroment
# When we use the here::here i do not have to insert the entier path the program knows where it is using the here (read setion 17.4)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
```


## Showing the data

Sent it to git - this is a small file.... for big files or sensetive files like human data put it in git ignore so it is not uploaded to the git. also large files will take up a lot of space!

```{r}
post_meal_data

glimpse(post_meal_data) # a different way to look at the data
```



# 18  Basic data visualization

## Plot one continous variable
They recomed that u only have one plot pr code chunck. it gives an option to add a figure lable to your html file... not sure i would do this tho.

```{r}
ggplot(post_meal_data, aes(x = BMI))


ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram(na.rm = FALSE) # na.rm is removing the missing valuse so set it to falls to get all data. (obs no missing data for this plot)
```


```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

##  18.5 🧑‍💻 Exercise: Plot a discrete variable



```{r}
#| fig-cap: "Group Distribution"
#| label: fig-Group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```


```{r}
#| fig-cap: "Glykemi Distribution"
#| label: fig-glykemi-bar
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

### Plotting more variables 
obs it cannot reander if it encounters an error 

```{r}
## | fig-cap: "Group and Glykemi Distribution"
## | label: fig-Group-Glykemi-bar-plot
# ggplot(post_meal_data, aes(x = Group, fill = glykemi)) +
# geom_bar()


# This gives a warning cuz fill wants a factor. so we have to convert glykemi to a factor.
```

Notes:
Use as.character() when:
- Data is free text
- Order doesn’t matter
- You don’t want R to treat values as categories

Use as.factor() when:
- Data represents categories
- Values repeat
- You’re doing statistical analysis or grouping
- Levels have meaning

```{r}
#| fig-cap: "Group and Glykemi Distribution"
#| label: fig-Group-Glykemi-bar-plot
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge()) # this position changes ti so that they stand next to eachouther and not on top
```


## Side by side plots
more plots in one figure

```{r fig-bmi-glycemia}
#| fig-cap: "BMI and glycemia, side by side"
#| fig-subcap:
#|  - "Distribution of BMI"
#|  - "Number of thoes with glycemia"
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```


## plotting two continous varables

```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose"
#| label: fig-bmi-auc-pg

ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() + # scatter plot
  geom_smooth() # trend curve
```

# 19  Collaborating with GitHub

This should be done in the consul:

```{r}

#is everything ok?
#usethis::gh_token_help()

#if the token is unset do this
#usethis::create_github_token()

#gitcreds::gitcreds_set() # give it the new token. 

# now check again if it says discovered all is well


```


now we will link this project to gitHub

```{r}
#usethis::use_github()

```





















































